Topics in Statistical Theory --- Example Sheet 1

### 2020-10-27 topics_statistical_theory/C2.tex
worked on some example sheets, updated msm and topics summaries (todo: dist and inv prob)

We can express the MSE in terms of convolutions: 
\begin{definition}
	Let $g_1, g_2 \colon \RR \to \RR$ be measurable. Then the \emph{convolution} of $g_1$ and $g_2$, denoted $g_1 * g_2$, is defined by
	\[
	(g_1 * g_2)(x) \ceq \int_\RR g_1(x - z) g(z) \dd{z}. 
	\]
\end{definition}
We can compute 
\begin{align*}
\Bias \hat f_n(x) &= \EE[\hat f_n(x)] - f(x) = \EE[K_h(x - X_1)] - f(x) = \int_\RR  K_h(x - z) f(z) \dd{z} \\
&= (K_h * f)(x) - f(x). 
\end{align*}

@@

\begin{proposition}
	Let $\hat{f_n}$ be the kernel density estimator with kernel $K$ and bandwidth $h > 0$ constructed from $X_1, \dotsc, X_n \iid f$. Then for any $x \in \RR, h > 0, n \in \NN$ we have
	\[
	\Var \hat f_n(x) \leq \frac{1}{nh} \norm{f}_\infty R(K). 
	\]
\end{proposition}

@@

>>>>

We can express the MSE in terms of convolutions: 
\begin{definition}
	Let $g_1, g_2 \colon \RR \to \RR$ be measurable. Then the \emph{convolution} of $g_1$ and $g_2$, denoted $g_1 * g_2$, is defined by
	\[
	(g_1 * g_2)(x) \ceq \int_\RR g_1(x - z) g(z) \dd{z}. 
	\]
\end{definition}
We can compute 
\begin{align}
\Bias \hat f_n(x) &= \EE[\hat f_n(x)] - f(x) = \EE[K_h(x - X_1)] - f(x) = \int_\RR  K_h(x - z) f(z) \dd{z} \nonumber \\
&= (K_h * f)(x) - f(x) \label{eq:bias_computation}. 
\end{align}

@@

\begin{proposition} \label{prop:variance_bound}
	Let $\hat{f_n}$ be the kernel density estimator with kernel $K$ and bandwidth $h > 0$ constructed from $X_1, \dotsc, X_n \iid f$. Then for any $x \in \RR, h > 0, n \in \NN$ we have
	\[
	\Var \hat f_n(x) \leq \frac{1}{nh} \norm{f}_\infty R(K). 
	\]
\end{proposition}

@@

Obtaining a bound on the bias is not at all straightforward: we wil need to introduce conditions on both the density $f$ and the kernel $k$. 
\begin{definition}
	Let $I \subseteq \RR$ be an interval, fix $\beta, L > 0$, and let $m \ceq \ceil{\beta} - 1$. A function $f \colon \RR \to \RR$ is said to belong to the \emph{Hölder class} $\HH(\beta, L)$ if $f$ is $m$ times differentiable on $I$ and
	\[
	\abs{f^{(m)}(x) - f^{(m)}(y)} \leq L \abs{x - y}^{\beta - m} \quad\text{for all $x, y \in I$}. 
	\]
	If $I$ is unspecified, we let $I = \RR$.

	The densities in $\HH(\beta, L)$ are denoted by
	\[
	\FF(\beta, L) \ceq \qty{f \in \HH(\beta, L) \mid f \geq 0 \text{ and } \int_\RR f \dd{x} = 1}. 
	\]
\end{definition}

\begin{definition}
	Fix $\ell \in \NN$. We say a kernel $k$ is of \emph{order} $\ell$ if $\int_\RR x^j k(x) \dd{x} =0 $ for $j = 1, \dotsc, \ell - 1$. 
\end{definition}

\begin{remark}
	Most kernels used in practice are of order 2. Note that a kernel of order $\geq 3$ cannot be nonnegative, since we have $\int_\RR x^2k(x) \dd{x} = 0$. Therefore, the kernels are not themselves densities and the corresponding kernel density estimate is not guaranteed to be a density. 
\end{remark}

\begin{proposition} \label{prop:bias_bound}
	Assume that $f \in \FF(\beta, L)$ and that $k$ is a kernel of order $\ell \ceq \ceil{\beta}$, and furthermore assume that
	\[
	\mu_\beta(k) \ceq \int_\RR \abs{u}^\beta \abs{k(u)} \dd{u} < \infty. 
	\]
	Then the kernel density estimate with bandwidth $h$ and kernel $k$ based on $X_1, \dotsc, X_n \sim f$ satisfies
	\[
	\abs{\Bias \hat f_n(x)} \leq \frac{L}{(\ell - 1)!} \mu_\beta(k) h^\beta \quad\text{for all $x \in \RR, h > 0, n \in \NN$}. 
	\]
\end{proposition}

\begin{proof}
	By \cref{eq:bias_computation}, we have 
	\[
	\Bias \hat f_n(x) = \frac1h \int_\RR k\qty(\frac{x - z}{h}) f(z) \dd{z} - f(x) = \int_\RR k(u)\qty( f(x - uh) - f(x)) \dd{x}. 
	\]
	By applying Taylor's theorem with the Lagrange remainder we obtain, with $m = \ceil{\beta} - 1$, that
	\[
	f(x - uh) - f(x) = \sum_{j=1}^{m-1} \frac{(-uh)^j}{j!} f^{(j)}(x) + \frac{(-uh)^m}{m!} f^{(m)}(x - \tau u h) \quad\text{for some $\tau \in [0, 1]$}. 
	\]
	Since $\int_\RR u^j k(u) \dd{u} = 0$ for all $j \leq m$, plugging the sum into the integral will give 0. Therefore, we find
	\[
	\Bias \hat f_n(x) =  \frac{(-h)^m}{m!} \int_\RR u^m k(u) f^{(m)}(x - \tau uh) \dd{u} = \frac{(-h)^m}{m!} \int_\RR u^m k(u) \qty[f^{(m)}(x - \tau uh)
- f^{(m)}(x)] \dd{u},	\]
	where the last inequality follows again from the fact that $k$ is of order $m + 1$.

	Now we use that $f \in \FF(\beta, L)$, and conclude
	\[
	\abs{\Bias \hat f_n(x)} \leq \frac{Lh^m}{m!} \int_\RR \abs{u}^m \abs{k(u)} \abs{\tau u h}^{\beta - m} \dd{u} \leq \frac{Lh^\beta}{m!} \int_\RR \abs{u}^\beta \abs{k(u)} \dd{u} = \frac{L}{(\ell - 1)!} \mu_\beta(k) h^\beta,
	\]	
	which concludes the proof. 
\end{proof}

Combining \cref{prop:variance_bound,prop:bias_bound}, we find that
\[
\MSE \hat f_n(x) \leq \frac{1}{nh} R(k) \norm{f}_\infty + \frac{L^2}{((\ell - 1)!)^2} \mu^2_\beta(k) h^{2\beta}. 
\]
By minimising this w.r.t.\ $h$, we find that the optimal $h$ is given by
\[
h_n^* = \qty(\frac{((\ell - 1)!)^2 \norm{f}_\infty R(k)}{2\beta L^2 \mu_\beta^2(k)})^{1/(2\beta + 1)} n^{-1/(2\beta + 1)}, 
\]
and the corresponding $\MSE$ is given by
\[
\MSE \hat f_n(x) \leq C n^{-2\beta/(2\beta + 1)},
\]
where $C$ is independent of $n$. This bound on the rate is slower than the $O(1/n)$ rate found in parametric problems, but such a rate is only obtained when the assumed model is correct.

\subsection{Bounds on the integrated variance and bias}
To bound the MISE, we will give bounds on the integrated variance and bias. 
\begin{proposition}
	Let $\hat{f_n}$ denote the kernel density estimate with bandwidth $h$ and kernel $k$ based on $X_1, \dotsc, X_n \iid P$ (where $P$ is a distribution on $\RR$). Then 
\end{proposition}
